---
title: "Modeling for Default"
subtitle: "Modeling default for loan applications with majority class baseline"
author: "Adam Bushman"
date: "2024-10-20"
format:
    html:
        embed-resources: true
execute:
    warning: false
---


# Approach

I'm focusing on modeling for `DEFAULT` using the majority class. This will establish a "baseline". We'll do so first with the imbalanced data and then with the balanced.

## Prep

Let's load a few libraries:

```{r}
library('tidyverse')
library('caret')
library('rsample')
library('pROC')
```

We'll load the cleaned, training data, both versions. We'll also just turn the target variable numeric for ease:

```{r}
imbal_data <- data.table::fread('D:/All Repos/home-credit-default-risk-group/data/application_train_clean.csv') |>
    as.data.frame() |>
    mutate(DEFAULT = ifelse(DEFAULT,1,0))

bal_data <- data.table::fread('D:/All Repos/home-credit-default-risk-group/data/application_train_smote.csv') |>
    as.data.frame() |>
    mutate(DEFAULT = ifelse(DEFAULT == 'Y',1,0))
```

## Calculation

All we have to do is assume the majority class is going to be the predicted value. Then we generate a confusion matrix and summary statistics. 

```{r}
table(imbal_data$DEFAULT) |> prop.table()
```

So `FALSE` is our majority classifier. Let's now assume all predictions are "FALSE" and generate a confusion matrix.

```{r}
imbal_data$pred <- 0
mat1 <- caret::confusionMatrix(factor(imbal_data$pred), factor(imbal_data$DEFAULT))
mat1
```

We can also calcualte the AUC value:

```{r}
roc_obj1 <- pROC::roc(imbal_data$DEFAULT, imbal_data$pred)
auc_val1 <- pROC::auc(roc_obj1)
auc_val1
```

Let's assemble all of this into a vector:

```{r}
performance1 <- c(
    c("model" = "Majority Class, Imbalanced"), 
    c("hyperparameters" = "None"), 
    mat1$overall[c("Accuracy")], 
    mat1$byClass[c("Precision", "Recall")], 
    c("AUC" = auc_val1)
)

performance1
```

Let's now do the same thing for the balanced data.

```{r}
bal_data$pred <- 0
mat2 <- caret::confusionMatrix(factor(bal_data$pred), factor(bal_data$DEFAULT))
mat2
```

```{r}
roc_obj2 <- pROC::roc(bal_data$DEFAULT, bal_data$pred)
auc_val2 <- pROC::auc(roc_obj2)
auc_val2
```

```{r}
performance2 <- c(
    c("model" = "Majority Class, Balanced"), 
    c("hyperparameters" = "None"), 
    mat2$overall[c("Accuracy")], 
    mat2$byClass[c("Precision", "Recall")], 
    c("AUC" = auc_val2)
)

performance2
```

Let's now turn this into a dataframe and write it to a CSV for pooling with future models.

```{r}
df <-
    data.frame(as.list(performance1)) |>
    rbind(
        data.frame(as.list(performance2))
    )

df
```

```{r}
#| eval: FALSE
#| include: FALSE
data.frame(as.list(df)) |> 
    write.csv('models/majority-class-baseline/model-results.csv', row.names = FALSE)
```